---
title: "Data Visualization code for Amr Proucts"
author: "mahmoud hassanen(@Mahmoud203) , Sarah Shebl (@Sarah50) , Abdullah Ibrahim Ali(@Abdullah108) , Kareem El Sayed Saad Abd El Gawad (Kareem_101) , Nada Esmael Sleim (Eddy27) "
date: "2024-09-14"        
output:
  cleanrmd::html_document_clean:
    theme: water-dark
---

# Introduction.

**Hi, this R markdown file for Data Visualization code.**

## R code.

first loading these Libraries

```{r}
#===================================library
library(tidyverse)
library(ggsci) 
library(ggplot2)
library(dplyr)

```

### Cloning The data

```{r}
#====================================loading data
data <- read.delim("https://raw.githubusercontent.com/HackBio-Internship/public_datasets/main/R/WHO_AMR_PRODUCTS_DATA.tsv", sep = "\t", header = TRUE)
```

### Cleaning the Data From '*N/A*' VALUES

```{r}
#====================================cleaning data 
AMR_Products <- data %>%
  filter(Active.against.priority.pathogens. !='N/A')%>%
  group_by(Product.type, Active.against.priority.pathogens.) 
```

### Explore The data

```{r, eval=FALSE}
view(AMR_Products)
dim(AMR_Products)
table(AMR_Products$Product.name)

```

### General Distribution of the Data

```{r}
# Distribution of Product Type
Product_type_summary<-AMR_Products%>%
  group_by(Product.type)%>%
  summarise(count=n())
Product_type_summary


ggplot(Product_type_summary, aes(x = Product.type, y = count, fill =Product.type)) +
  geom_bar(stat = 'identity',width = 0.3,position = 'dodge') +
  scale_fill_manual(values = c("#4F81BD", "#C0504D"))  +
  theme_minimal()+xlab('')+
  theme(text = element_text(size = 9),
        plot.title = element_text(hjust =0.5,face = 'bold' ),legend.position = "bottom")

```

### create Drugs effective against The Pathogen Data

```{r, eval=FALSE}
#Test the product activity
product_activity<-AMR_Products%>%group_by(Product.type,Active.against.priority.pathogens.)%>%
  summarise(Count=n())
product_activity
```

### Visualize the data

```{r}
product_activity<-AMR_Products%>%group_by(Product.type,Active.against.priority.pathogens.)%>%
  summarise(Count=n())
product_activity
# Create the stacked bar chart that shows distribution of Product Type with activity status
ggplot( product_activity, aes(x = Product.type, y = Count, fill = Active.against.priority.pathogens.)) +
  geom_bar(stat = "identity", width = 0.8) +  
  scale_fill_manual(values = c("Yes" = "#FF5733", "No" = "grey", "Possibly" = "#3357FF")) +
  labs(
    x = "Product Type",
    y = "Count",
    fill = "Activity Status"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

### summarize the Data

```{r}
Route_Activity_relation <- AMR_Products %>%
  group_by(Route.of.administration, Active.against.priority.pathogens.) %>%
  summarise(Count = n(), .groups = 'drop')
Route_Activity_relation
```

### Relation between Route Route.of administration and activity status

```{r}
ggplot_plot <- ggplot(Route_Activity_relation, aes(x = Route.of.administration, y = Count, color = Active.against.priority.pathogens.)) +
  geom_point(size = 3) +
  scale_color_manual(values = c("Yes" = "#FF5733", "No" = "grey", "Possibly" = "#3357FF")) +
  labs(title = "Relationship Between Route of Administration and Drug Effectiveness",
       x = "Route of Administration",
       y = "Number of Products",
       color = "Effectiveness Against Priority Pathogens") +
  theme_minimal() +
  theme(legend.position = "bottom", legend.title.position = "bottom")
ggplot_plot
```

# Antibiotics

## create Antibiotics data

```{r}
#================================Antibiotics data
Antibiotics <- data %>% filter(Product.type == "Antibiotics")

```

## what is the R.D phase of antibiotic products

```{r}
ggplot(Antibiotics,mapping  = aes(R.D.phase,fill=Product.name))+geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(50))+
  theme(axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 10,angle = 45, hjust = 1),
        axis.ticks.y = element_blank())
```

## what is The Route.of.administration of Antibiotics data

```{r}
#what is the route of adminstrtion of the products
ggplot(Antibiotics,mapping = aes(Route.of.administration,fill=Product.name))+geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(46))+
  theme(axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 10,angle = 45, hjust = 1),
        axis.text.y = element_text(size = 10)
  )
```

### loading Antibiotic activity data

this is The Antibiotics That has show Result against priority Pathogen .

```{r}
ggplot(Antibiotics,mapping =aes(Active.against.priority.pathogens.,fill=Product.name))+
  geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(50))
ggplot

```

### Visualize the data

```{r}
#loading Antibiotics data that had been shown result after testing 
Antibiotics <- AMR_Products %>% filter(Product.type == "Antibiotics")

```

# Non Traditional categories.

create ***Non Traditional categories*** data.

```{r}
# non traditional data categories
Non.traditional = AMR_Products%>%filter(Product.type=="Non-traditional")
colnames(Non.traditional)

```

## Visualize the data

```{r}
ggplot(Non.traditional, aes(x = "", y = Non.traditionals.categories, fill = factor(Non.traditionals.categories))) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_discrete(name = "Non.traditionals.categories") +
  labs(title = "Pie Chart of Non.traditionals.categories") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom",legend.title.position = "top")


```

## Which category has effect and on which pathogens

```{r}
# Create a simple dot plot 
ggplot(Non.traditional, aes(x = Non.traditionals.categories, y = Pathogen.name)) +
  geom_point(aes(color = Pathogen.name), size = 3, position = position_jitter(width = 0.2, height = 0.2)) +
  labs(title = "Effect of Products on Pathogens",
       x = "Non.traditionals.categories",
       y = "Pathogen Name",
       color = "Pathogen Name") +
  theme(legend.position = "bottom")
```

# Bacteriophages and phage derived enzymes

## create Bacteriophages and phage derived enzymes Data

```{r}
Bacteriophages_and_phage_derived_enzymes_data <- data %>%
  filter(Non.traditionals.categories == "Bacteriophages and phage-derived enzymes")


```

### what is the route of administration

```{r}
ggplot(Bacteriophages_and_phage_derived_enzymes_data,mapping = aes(Route.of.administration,fill=Product.name))+
  geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))

```

### what is R.D phase of this products

```{r}
ggplot(Bacteriophages_and_phage_derived_enzymes_data,mapping = aes(R.D.phase,fill=Product.name))+
  geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))

```

## products of Bacteriophages and phage derived enzymes against pathogens.

### loading data

```{r}
Bacteriophages_and_phage_derived_enzymes_data <- AMR_Products %>% filter(Non.traditionals.categories == "Bacteriophages and phage-derived enzymes")

```

###visualize data

```{r}
#visualize data
ggplot(Bacteriophages_and_phage_derived_enzymes_data,mapping = aes(Active.against.priority.pathogens.,fill=Product.name))+geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))+
  theme(axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 10,angle = 45, hjust = 1),
        axis.text.y = element_text(size = 10)
  )
```

#Immunomodulating agents data

## create Immunomodulating agents data

```{r}
Immunomodulating_agents_data <- data %>% filter(Non.traditionals.categories == "Immunomodulating agents")

```

## Products Used in Immunomodulating agents data

```{r}
ggplot(Immunomodulating_agents_data,mapping = aes(Antibacterial.class,fill=Product.name))+
  geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))

```

## what is R.D phase of this products

```{r}
ggplot(Immunomodulating_agents_data,mapping = aes(R.D.phase,fill=Product.name))+
  geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))

```

## what is the route of administration

```{r}
ggplot(Immunomodulating_agents_data,mapping = aes(Route.of.administration,fill=Product.name))+
  geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))

```

## Immunomodulating_agents_activity_data that have been Tested in activity against priority pathogen data

### create data

```{r}
Immunomodulating_agents_activity_data <- AMR_Products %>% filter(Non.traditionals.categories == "Immunomodulating agents")

```

### Application on different pathogens

```{r}
Immunomodulating_agents <- Non.traditional %>%
  filter(Product.name %in% c("Rhu-pGSN", "AB103"))
Immunomodulating_agents
```

### visualize it

```{r}
ggplot(Immunomodulating_agents, aes(x = Product.name, y = Pathogen.name)) +
  geom_point(aes(color = Pathogen.name), size = 3, position = position_jitter(width = 0.2, height = 0.2)) +
  labs(title = "Effect of Rhu-pGSN and AB103 on Pathogens",
       x = "Product Name",
       y = "Pathogen Name",
       color = "Pathogen Name") +
  theme(
    text = element_text(size = 16),  # Set base font size for all text
    plot.title = element_text(size = 20, face = "bold"),  
    axis.title.x = element_text(size = 18),               
    axis.title.y = element_text(size = 18),               
    axis.text.x = element_text(size = 14, angle = 25, hjust = 1),  
    axis.text.y = element_text(size = 14),                 
    legend.text = element_text(size = 14),                 
    legend.title = element_text(size = 16),                
    legend.position = "bottom"                             
  )
```

# Microbiome-modulating agents

## Create Microbiome-modulating agents data

```{r}
Microbiome_modulating_agents <- data %>% filter(Non.traditionals.categories == "Microbiome-modulating agents")

```

## what is the anti bacterial class that have been used in Microbiome-modulating agents.

```{r}
ggplot(Microbiome_modulating_agents,mapping = aes(Antibacterial.class,fill=Product.name))+
  geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))

```

## what is R.D phase of this products.

```{r}
ggplot(Microbiome_modulating_agents,mapping = aes(R.D.phase,fill=Product.name))+
  geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))

```

## what pathogens had been tried against products

```{r}
ggplot(Microbiome_modulating_agents,mapping = aes(Pathogen.name,fill=Product.name))+
  geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))

```

## what is the route of administration

```{r}
ggplot(Microbiome_modulating_agents,mapping = aes(Route.of.administration,fill=Product.name))+
  geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))

```

## LOADING Data of microbe modulating agents filtrate from *N/A*

```{r}
microbe_modulating_agents_data <- AMR_Products %>% filter(Non.traditionals.categories == "Microbiome-modulating agents")

```

## Visualize data

```{r}
ggplot(microbe_modulating_agents_data,mapping = aes(Active.against.priority.pathogens.,fill=Product.name))+geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))+
  theme(axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 10,angle = 45, hjust = 1),
        axis.text.y = element_text(size = 10)
  )
```

# Miscellaneous

## loading Miscellaneous data

```{r}
Miscellaneous_data <- data %>% filter(Non.traditionals.categories == "Miscellaneous")
```

## what is the anti bacterial class that have been used in Miscellaneous.

```{r}
ggplot(Miscellaneous_data,mapping = aes(Antibacterial.class,fill=Product.name))+geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))+
  theme(axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 10,angle = 45, hjust = 1),
        axis.text.y = element_text(size = 10))+coord_flip()

```

## what is R.D phase of this products

```{r}
ggplot(Miscellaneous_data,mapping = aes(R.D.phase,fill=Product.name))+geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))+
  theme(axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 10,angle = 45, hjust = 1),
        axis.text.y = element_text(size = 10)
  )
```

## what is the route of administration

```{r}
ggplot(Miscellaneous_data,mapping = aes(Route.of.administration,fill=Product.name))+geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))+
  theme(axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 10,angle = 45, hjust = 1),
        axis.text.y = element_text(size = 10)
  )
```

##products of Miscellaneous against pathogens.

## loading_data

```{r}
Miscellaneous_activity_data <- AMR_Products %>% filter(Non.traditionals.categories == "Miscellaneous")

```

## visualize data

```{r}
ggplot(Miscellaneous_activity_data,mapping = aes(Active.against.priority.pathogens.,fill=Product.name))+geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))+
  theme(axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 10,angle = 45, hjust = 1),
        axis.text.y = element_text(size = 10))

```

# Antibodies data.

## Loading Antibodies data.

```{r}
antibodies_data <- data %>% filter(Non.traditionals.categories == "Antibodies")

```

## Visualize Antibodies data.

the antibacterial class of Antibodies

```{r}
ggplot(antibodies_data,mapping = aes(Antibacterial.class,fill=Product.name))+geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))+
  theme(axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 10,angle = 45, hjust = 1),
        axis.text.y = element_text(size = 10))+coord_flip()

```

## what is R.D phase of this products ?

```{r}
ggplot(antibodies_data,mapping = aes(R.D.phase,fill=Product.name))+
  geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))

```

## what pathogens had been tried against products

```{r}
ggplot(antibodies_data,mapping = aes(Pathogen.name,fill=Product.name))+geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))+
  theme(axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 10,angle = 45, hjust = 1),
        axis.text.y = element_text(size = 10)
  )
```

##What is the route of administration ?

```{r}
ggplot(antibodies_data,mapping = aes(Route.of.administration,fill=Product.name))+
  geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))

```

## products of antibodies against pathogens.

## Loading data and Visualize it

```{r}
#loading_data
antibodies_activity_data <- AMR_Products %>% filter(Non.traditionals.categories == "Antibodies")
#visualize data
ggplot(antibodies_activity_data,mapping = aes(Active.against.priority.pathogens.,fill=Product.name))+
  geom_bar()+theme_light()+scale_fill_manual(values = pal_igv()(20))

```

That is all the code that has been used up **we also created** :

1.  Notion [Report](https://vigorous-dahlia-f5d.notion.site/AMR-New-Products-Deep-Analysis-38b6b444f6224a15a3afd647f7e6c18f?pvs=4)

2.  Info graphic[ summary](https://drive.google.com/file/d/1xnLhcd_LJfnsJomjtbc3gQtF7GAVvwgg/view)
